---
title: "Assingment 2"
output: html_document
date: "2023-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
library(corrgram)
library(visdat)
library(forecast)
library(tidytext)
library(tidyverse)
library(janeaustenr)
library(stringr)
library(wordcloud2)
library(reshape2)
library(pls)
library(ggplot2)
library(devtools)
library(car)
#library(tabplot)
library(textdata)
library(seriation)
library(visNetwork)
library(leaflet)
library(shinycssloaders)
library(DT)
library(plotly)

library(data.table)
library(jstable)
library(GGally)
library(summarytools)
library(ggstatsplot)
library(naniar)
library(rpart)
library(rpart.plot)
library(caret)
library(glmnet)
library(recipes)
library(skimr) 
library(tidymodels) 
library(caret)
library(glmnet)
```

## Import data

Missing placeholders : "NA", -99, "\--",

```{r}
dat <- read.csv("Ass2Data.csv", header = TRUE, na.strings =
c('NA','-99', '--'), stringsAsFactors = TRUE)
```

```{r}
summary(dat)
```

5\. Identify any categorical missing values that are "Not Applicable" and create new levels for these values using something like:

```{r}
dat_2 = dat

dat_2$POLITICS <- as.character(dat_2$POLITICS) #convert away from factor
dat_2$POLITICS[is.na(dat_2$POLITICS)] <- "NONE"
dat_2$POLITICS <- as.factor(dat_2$POLITICS) # convert back to factor

dat_2$HEALTHCARE_BASIS <- as.character(dat_2$HEALTHCARE_BASIS) #convert away from factor
dat_2$HEALTHCARE_BASIS[is.na(dat_2$HEALTHCARE_BASIS)] <- "NONE"
dat_2$HEALTHCARE_BASIS <- as.factor(dat_2$HEALTHCARE_BASIS) # convert back to factor
```

6.  SHADOW VARIABLES

Identify any numeric missing values that are "Not Applicable" and create new levels for these values

```{r}
# dat_2$POPULATION_shadow <- as.numeric(is.na(dat_2$POPULATION)) 
# #dat_2$POPULATION[is.na(dat_2$POPULATION)] <- 0 
# 
# dat_2$AGE25_PROPTN_shadow <- as.numeric(is.na(dat_2$AGE25_PROPTN))
# #dat_2$AGE25_PROPTN[is.na(dat_2$AGE25_PROPTN)] <- 0 
# 
# dat_2$AGE50_PROPTN_shadow <- as.numeric(is.na(dat_2$AGE50_PROPTN)) 
# #dat_2$AGE50_PROPTN[is.na(dat_2$AGE50_PROPTN)] <- 0 
# 
# dat_2$AGE_MEDIAN_shadow <- as.numeric(is.na(dat_2$AGE_MEDIAN)) 
# #dat_2$AGE_MEDIAN[is.na(dat_2$AGE_MEDIAN)] <- 0 
# 
# dat_2$POP_DENSITY_shadow <- as.numeric(is.na(dat_2$POP_DENSITY)) 
# #dat_2$POP_DENSITY[is.na(dat_2$POP_DENSITY)] <- 0 
# 
# dat_2$GDP_shadow <- as.numeric(is.na(dat_2$GDP)) 
# #dat_2$GDP[is.na(dat_2$GDP)] <- 0 
# 
# dat_2$INFANT_MORT_shadow <- as.numeric(is.na(dat_2$INFANT_MORT)) 
# #dat_2$INFANT_MORT[is.na(dat_2$INFANT_MORT)] <- 0 
# 
# dat_2$DOCS_shadow <- as.numeric(is.na(dat_2$DOCS)) 
# #dat_2$DOCS[is.na(dat_2$DOCS)] <- 0 
# 
# dat_2$VAX_RATE_shadow <- as.numeric(is.na(dat_2$VAX_RATE)) 
# #dat_2$VAX_RATE[is.na(dat_2$VAX_RATE)] <- 0 
# 
# dat_2$HEALTHCARE_COST_shadow <- as.numeric(is.na(dat_2$HEALTHCARE_COST)) 
# #dat_2$HEALTHCARE_COST[is.na(dat_2$HEALTHCARE_COST)] <- 0 
# 
# dat_2$DEATH_RATE_shadow <- as.numeric(is.na(dat_2$DEATH_RATE)) 
# #dat_2$DEATH_RATE[is.na(dat_2$DEATH_RATE)] <- 0 
```

Data clean : Variables

```{r}
pMiss <- function(x){ sum(is.na(x)) / length(x) * 100 }

threshold <- 50
d <- dat_2
cRatio <- apply(X = d, MARGIN = 2, FUN = pMiss) # run pMiss for each column of the data frame
cat("Variables to remove:", paste(colnames(d)[cRatio > threshold], collapse = ","))
```

Missingness threshold and list rows to remove

```{r}
threshold <- 50

rRatio <- apply(X = d, MARGIN = 1, FUN = pMiss)  # run pMiss for each row of the df 
cat("Observations to remove (First 50) :", paste(head(rownames(d)[rRatio > threshold], n = 50), collapse = ", "))
```

Clean both Vars and Obs

```{r}
cat("Original dimensions: ",dim(d))

var_miss_threshold = 100
obs_miss_threshold = 100

# remove variables first
cRatio <- apply(d,2,pMiss)
d <- d[, cRatio < var_miss_threshold]

# then observations
rRatio <- apply(d,1,pMiss)
d <- d[rRatio < obs_miss_threshold, ]

cat("Final data dimensions are: ",dim(d))

```

Outlier cleaning strategy

**Interquartile range method:**

only be interested in identifying outliers in one column of a data frame. For example, suppose we only want to remove rows that have an outlier in column 'A' of our data frame.

11.Create a test - train split

```{r}
train <- d[d$OBS_TYPE == 'Train',]
test <- d[d$OBS_TYPE == 'Test',]
```

Recipe based processing pipeline:

```{r}
recipeA <- recipes::recipe(DEATH_RATE ~., data = train) %>%
      update_role("CODE", new_role = 'id') %>% #id is not a predictor
      update_role('OBS_TYPE', new_role = 'split') %>% #obs_type is not a predictor
      step_impute_knn(all_predictors(), neighbors = 5) %>%
      step_rm('OBS_TYPE') %>%
      step_rm("CODE") %>%
      step_center(all_numeric(), -has_role("outcome"))%>%
      step_scale(all_numeric(), -has_role("outcome")) %>%
      step_dummy(all_predictors(), -all_numeric())

summary(recipeA)

prepped_recipeA <- prep(recipeA, training = train)
prepped_recipeA

covid_train_baked <- bake(prepped_recipeA, train) 
covid_train_baked

covid_test_baked <- bake(prepped_recipeA, test)
covid_test_baked
```

```{r}
fit.control <- trainControl(method = "repeatedcv", number = 5, repeats = 10)

set.seed(123)
model <- caret::train(
              recipeA, 
              data = train, 
              tuneGrid = expand.grid(
                alpha = 0:1,
                lambda = seq(0.0001, 1, length = 20)
              ),
              method = 'glmnet',  
              trControl = fit.control)
model
model.res = resid(model)
model.res
```

```{r}
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}

model$modelType
get_best_result(model)
print (max(model[["results"]][["RMSE"]]))
```

```{r}
plot(model)
```

RMSE

```{r}
calc_rmse = function(actual, predicted) {
  sqrt(mean((actual - predicted) ^ 2))
}

calc_rmse(actual = test$DEATH_RATE,
          predicted = predict(model, test))
```

```{r}
test$Prediction <- predict(model, newdata = test)
caret::RMSE(test$Prediction, test$DEATH_RATE, na.rm = TRUE)
```

PREDICTIONS

```{r}
rang <- range(c(test$DEATH_RATE, test$Prediction))
ggplot(data = test) +
  geom_point(mapping = aes(x = Prediction, y = DEATH_RATE)) +
  geom_abline(slope = 1, col = "blue") +
  labs(title = "Death Rate predictions of covid test data", y = "predicted", x = "actual") +
  coord_fixed(ratio = 1, xlim = rang, ylim = rang, expand = TRUE)
```

```{r}
summary (model)
```

```{r}
# Calculate residuals for test and train data
test$Residuals <- residuals(model$finalModel, newdata = test)
train$Residuals <- residuals(model$finalModel, newdata = train)
```

```{r}
best_model = model$finalModel

#different way of getting residuals - assigning to variables. 
# Get residuals for train data
train_resid <- resid(best_model)

# Get residuals for test data
test$Prediction <- predict(best_model, newx = test)
test_resid <- test$DEATH_RATE - test$Prediction
```

```{r}
# Calculate residuals for the test and train data
test$Residuals <- test_resid
train$Residuals <- residuals(model)

# Create a data frame with residuals and data set indicator
data <- rbind(
  data.frame(Data = "Test", Residuals = test$Residuals),
  data.frame(Data = "Train", Residuals = train$Residuals)
)

# Create boxplot with ggplot2
library(ggplot2)
p <- ggplot(data, aes(x = Data, y = Residuals)) + 
  geom_boxplot()

# Label outliers
outliers <- boxplot(data$Residuals, plot = FALSE)$out
p + geom_text(aes(x = Data, y = outliers, label = round(outliers, 2)), 
              hjust = -0.2, vjust = 0.5)
```
